---
title: "Marker analysis of 6d TB Reference and 4d TB Treatment"
format: 
  html:
    toc: true
    toc-location: right
    smooth-scroll: true
    code-fold: true
    code-summary: "Show the code"
execute:
  cache: false
  warning: FALSE
  message: FALSE
---


```{r}
library(readxl)
library(janitor)
library(tidyverse)
```


```{r data}
final_4d_tb_Reference <- read_excel("4d_TB_reference vs treatment timer_SM_for stats.xlsx", 
    sheet = "4d_TB_reference pattern_Final")

final_4d_tb_Treatment <- read_excel("4d_TB_reference vs treatment timer_SM_for stats.xlsx", 
    sheet = "FINAL_4d_tb DURING TREATMENT")
```

```{r}
final_4d_tb_Reference |> clean_names() |> select(sample_position, x5_0_sec:x1_8_sec) |> 
  pivot_longer(
    names_to = "time", values_to = "marker value", x5_0_sec:x1_8_sec) |> 
  drop_na() |> mutate(time = str_replace_all(time, "x", ""),
                      type = "4d TB Reference") -> df1
```

#### 4d TB Reference Data overview

```{r}
df1 |> ggplot(aes(sample_position, `marker value`, col=time)) +
  geom_boxplot(outlier.shape = NA) + 
  labs(title = "4d TB Reference data overview")
```

#### Distribution

```{r}
df1 |> ggplot(aes(sample_position, `marker value`, col = time)) +
  geom_line() +
  labs(title = "Distribution of 4d TB Reference based on time point")
```

In every time point marker values follows the polynomial distribution for more clarity plot separately.

```{r}
df1 |> ggplot(aes(sample_position, `marker value`, col = type)) +
  geom_line() + facet_wrap(~time)
```

### 4d TB Treatment
```{r}
final_4d_tb_Treatment |> clean_names() |> 
  select(sample_position, x2_0_sec:x5_0_sec) |> 
  pivot_longer(
    names_to = "time", values_to = "marker value", x2_0_sec:x5_0_sec) |> 
  drop_na() |> mutate(time = str_replace_all(time, "x", ""),
                      type = "6d TB Treatment") -> df2
```


#### Data overview

```{r}
df2 |> ggplot(aes(sample_position, `marker value`, col=time)) +
  geom_boxplot(outlier.shape = NA) + 
  labs(title = "6d TB Reference")
```

#### Distribution

```{r}
df2 |> ggplot(aes(sample_position, `marker value`, col = time)) +
  geom_line() +
  labs(title = "Distribution of 6d TB Reference based on time point")
```

In every time point marker values follows the polynomial distribution for more clarity plot separately.

```{r}
df2 |> ggplot(aes(sample_position, `marker value`, col = type)) +
  geom_line() + facet_wrap(~time)
```


#### Combine both

```{r}
newdf <- rbind(df1, df2)
newdf |>  
  ggplot(aes(sample_position, `marker value`, col = type)) +
  geom_line() + facet_wrap(~time)
```
Recoding marker time of this two data sets are not same

<!-- #### Polynomial regression -->

<!-- ```{r} -->
<!-- results_df <- data.frame() -->
<!-- coefficients_df <- data.frame() -->

<!-- # Loop through all unique combinations of 'time' and 'type' -->
<!-- for (time_val in unique(newdf$time)) { -->
<!--   for (type_val in unique(newdf$type)) { -->
<!--     # Subset the data for the current combination -->
<!--     subset_data <- subset(newdf, time == time_val & type == type_val) -->

<!--     # Perform polynomial regression -->
<!--     model <- lm(`marker value` ~ poly(sample_position, degree = 4), data = subset_data) -->

<!--     coef_names <- c("Intercept", "x1", "x2", "x3", "x4") -->
<!--     coefficients_subset <- data.frame( -->
<!--       type = type_val, -->
<!--       time = time_val, -->
<!--       coefficients = setNames(as.list(round(coef(model),3)), coef_names)  #, coef_names -->
<!--     ) -->
<!--     coefficients_df <- bind_rows(coefficients_df, coefficients_subset) -->

<!--     # Predict values -->
<!--     predicted_values <- predict(model, newdata = subset_data) -->

<!--     # Create a data frame with results -->
<!--     result_subset <- data.frame( -->
<!--       time = time_val, -->
<!--       type = type_val, -->
<!--       sample_position = subset_data$sample_position, -->
<!--       marker_value = subset_data$`marker value`, -->
<!--       predicted = predicted_values -->
<!--     ) -->

<!--     # Append the results to the main data frame -->
<!--     results_df <- bind_rows(results_df, result_subset) -->
<!--   } -->
<!-- } -->

<!-- colnames(coefficients_df) <- c("type", "time", "Intercept", "x1", "x2", "x3", "x4") -->
<!-- coefficients_df -->
<!-- ``` -->

<!-- #### Graphical representation -->
<!-- ```{r} -->
<!-- ggplot(results_df, aes(y = marker_value, x = sample_position)) + -->
<!--   geom_line(col = "black") + -->
<!--   geom_line(aes(y = predicted), linetype = "dashed", col = "red") + -->
<!--   facet_grid(type ~ time, scales = "free_y")  + -->
<!--   labs(title = "Polynomial Regression", x = "Marker Value", y = "Sample Position") -->
<!-- ``` -->

<!-- only 2.0 sec 4 dot maker and 6 dot marker coefficient magnitude not same rest of that are same so this two marker time point marker values magnitude are same. -->